Incentivizing Exploration by Heterogeneous Users

### 2018-02-17 lower_bound.tex

We saw in Corollary~\ref{cor:constant-regret} that \bccomment{David, could you put the right reference here?}
when agent preferences for their best arm are sufficiently clear,
in the sense that $\TieDensity = 0$, the regret of
Algorithm~\ref{alg:basic-incentivizing} is bounded by a constant.
One may conjecture that this should hold more generally,
in that the regret of the (fewer and fewer) agents on the boundary
between close arms should go to 0, while their fraction also goes to 0.
In this section, we establish a lower bound,
showing that even in very simple settings, a (logarithmic) dependence
on $T$ is typicaly unavoidable for \emph{any} incentivization strategy.

>>>>

We saw in Theorem~\ref{rst:discrete} that 
when agent preferences for their best arm are sufficiently clear,
in the sense that $\TieDensity = 0$, the regret of
Algorithm~\ref{alg:basic-incentivizing} is bounded by a constant.
One may conjecture that this should hold more generally,
in that the regret of the (fewer and fewer) agents on the boundary
between close arms should go to 0, while their fraction also goes to 0.
In this section, we establish a lower bound,
showing that even in very simple settings, a (logarithmic) dependence
on $T$ is typicaly unavoidable for \emph{any} incentivization strategy.

### 2018-02-17 problem_setting.tex

We consider a MAB setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms can be identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, but unknown to the agents
or the principal.
Whenever arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero
Gaussian%
\footnote{For simplicity of notation, we assume that the variance
  $\sigma^2$ is uniform across time steps. This is not material for
  the analysis.}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
\dkcomment{Clarify if we can handle more general noise.}
Here, $I_d$ denotes the $d \times d$ identity matrix.

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his%
\footnote{We use male pronouns to refer to users and female pronouns
  to refer to the principal. 
  \dkcomment{Feel free to switch if you prefer it the other way.}}
\emph{type}) $\AgentV{t} \in \R^d$ is drawn from a known distribution
\AgentDist.
Depending on context, we will identify a agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.

@@

Since each user only pulls an arm once, users are \emph{myopic}:
in the absence of incentives, user $t$ will pull an arm from
$\argmax_i \AgentV{t} \cdot \ArmEV{t}{i}$.
To incentivize users to explore more, the principal can offer
\emph{payments} $\Pay{t}{i}$ to user $t$ for pulling arm $i$.
Then, user $t$ will pull an\footnote{We assume that ties are broken in
  favor of an arm with largest payment \Pay{t}{i}; this assumption is
  only for notational convenience, and can of course be avoided by
  raising payments infinitesimally.} arm $i$ maximizing
$\argmax_i (\Pay{t}{i} + \AgentV{t} \cdot \ArmEV{t}{i})$.
The principal cannot observe \AgentV{t},
and only knows the distribution \AgentDist from which it is drawn.
Her goal is to achieve a good tradeoff between the cumulative
\emph{regret} experienced by all users up to time $T$,
and the total \emph{payment} she makes to the users.

We define the regret at time $t$ as
$\Regret{t} = (\max_{i} \AgentV{t} \cdot \ArmV{i}) - \AgentV{t} \cdot \ArmV{\Pull{t}}$,
and the cumulative regret up to time $T$ as
$\TotalRegret{T} = \sum_{t=1}^{T} \Regret{t}$.
Similarly, $\PayA{t} = \Pay{t}{\Pull{t}}$ is the actual incentive
payment at time $t$,
and the cumulative payment up to time $T$ is
$\TotalPay{T} = \sum_{t=1}^{T} \PayA{t}$.
More formally, the principal's goal is to find a policy
\POLICY for offering payments under which both the cumulative expected
regret
$\Expect[\POLICY]{\TotalRegret{T}}$ the cumulative expected payment
$\Expect[\POLICY]{\TotalPay{T}}$ are small.

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV;
\dkdelete{breaking ties uniformly at random}.
\dkcomment{What would it mean to break ties at random in a
  definition?}

\dkcomment{Do we need this definition here?}
We call time $t_{n}=\min_{t}\{\forall i, \NumPull{t}{i} \geq n\}$ the
\emph{starting point of the $n^{th}$ round}.
We call the set of times $[t_{n}, t_{n+1})$ the
\emph{$n^{th}$ round}.

\dkdelete{This behavior may be recovered if agents are Bayesian and
  share a common non-informative prior distribution that is constant
  over $\mathbb{R}^m$ and know $\sigma^2$.  In this case, the
  posterior distribution on $u_{i}$ at time $t$ is multivariate normal
  with mean $u_{i,t}$, and the expected value of $\theta_t \cdot u_i$
  under this posterior conditioned on $\theta_t$ is $\theta_t \cdot
  u_{i,t}$ (see Equation 2.13 in Section 2.5, \cite{Ge04}).
  Alternatively, one may simply take our assumption that agents use
  the average as their estimate of an attribute value directly without
  such a Bayesian justification.}
\dkcomment{I wonder what's the best way to discuss this.}

@@

Let $\FirstTwo{i}{i'} = \Set{\AgV}{\Best{\AgV} = i, \Second{\AgV} = i'}$
be the set of agent attribute vectors whose best arm is $i$ and
second-best arm is $i'$.
Let $F_{i,i'}$ be the marginal cumulative density function
(or cumulative mass function if $\AgentDist(\cdot)$ is a discrete distribution)
of $(\ArmV{i}-\ArmV{i'}) \cdot \AgV$,
conditioned on $\AgV \in \Omega_{i,i^{'}}$.
In words, $F_{i,i'}$ is the distribution of the \emph{strength} of the
preference of a random agent for arm $i$ over arm $i'$, given that
$i$ and $i'$ are the agent's first and second choices.

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exists a constant \TieDensity such that
for all pairs $i,i'$ of arms and all $z \in \R^+$,
we have $F_{i,i'}(z) \leq \TieDensity \cdot z$.
\end{assumption}

>>>>

We consider a multi-armed bandit setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms can be identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, and unknown to the agents
or the principal.
Whenever arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero independent Gaussian%
\footnote{For simplicity of notation, we assume that the variance
  $\sigma^2$ is uniform across time steps.
  This is not material for the analysis.
\dkedit{In fact, the analysis extends straightforwardly to any
  mean-zero sub-Gaussian noise.} \dkcomment{Is this true?}}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
Here, $I_d$ denotes the $d \times d$ identity matrix.

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.%
\footnote{\dkedit{Some of our results generalize to other forms of
    agent rewards.}}

@@

Since each user only pulls an arm once, users are \emph{myopic}:
in the absence of incentives, user $t$ will pull an arm from
$\argmax_i \AgentV{t} \cdot \ArmEV{t}{i}$.
To incentivize users to explore more, the principal can offer
\emph{payments} $\Pay{t}{i}$ to user $t$ for pulling arm $i$.
Then, user $t$ will pull an\footnote{We assume that ties are broken in
  favor of an arm with largest payment \Pay{t}{i}.}
% ; this assumption is
%   only for notational convenience, and can of course be avoided by
%   raising payments infinitesimally.
%   \pfcomment{In the finite preference setting this is clear, but it's a little bit delicate when there are continuum preferences, since regardless of the payment we offer there may always be support within our preference distribution for a $\theta$ that causes a tie.  In this continuum setting it should instead be possible to perturb infinitessimally so that we avoid ties with probability 1.}  
% }
arm $i$ maximizing $\argmax_i (\Pay{t}{i} + \AgentV{t} \cdot \ArmEV{t}{i})$.
The principal cannot observe \AgentV{t},
and only knows the distribution \AgentDist from which it is drawn.
Her goal is to achieve a good tradeoff between the cumulative
\emph{regret} experienced by all users up to time $T$,
and the total \emph{payment} she makes to the users.

We define the regret at time $t$ as
$\Regret{t} = (\max_{i} \AgentV{t} \cdot \ArmV{i}) - \AgentV{t} \cdot \ArmV{\Pull{t}}$,
and the cumulative regret up to time $T$ as
$\TotalRegret{T} = \sum_{t=1}^{T} \Regret{t}$.
Similarly, $\PayA{t} = \Pay{t}{\Pull{t}}$ is the actual incentive
payment at time $t$,
and the cumulative payment up to time $T$ is
$\TotalPay{T} = \sum_{t=1}^{T} \PayA{t}$.
More formally, the principal's goal is to find a policy
\POLICY for offering payments under which both the cumulative expected
regret
$\Expect{\TotalRegret{T}}$ the cumulative expected payment
$\Expect{\TotalPay{T}}$ are small.
\dkcomment{Deleted the \POLICY subscript for the expectation, as the
  expectation is also over the entire random history.}

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV,
\dkedit{breaking ties arbitrarily (but consistently)}.

% \dkdelete{This behavior may be recovered if agents are Bayesian and
%   share a common non-informative prior distribution that is constant
%   over $\mathbb{R}^m$ and know $\sigma^2$.  In this case, the
%   posterior distribution on $u_{i}$ at time $t$ is multivariate normal
%   with mean $u_{i,t}$, and the expected value of $\theta_t \cdot u_i$
%   under this posterior conditioned on $\theta_t$ is $\theta_t \cdot
%   u_{i,t}$ (see Equation 2.13 in Section 2.5, \cite{Ge04}).
%   Alternatively, one may simply take our assumption that agents use
%   the average as their estimate of an attribute value directly without
%   such a Bayesian justification.}
% \dkcomment{I wonder what's the best way to discuss this.}

@@

%Let $\FirstTwo{i}{i'} = \Set{\AgV}{\Best{\AgV} = i, \Second{\AgV} = i'}$
%be the set of agent attribute vectors whose best arm is $i$ and
%second-best arm is $i'$.
% Let $F_{i,i'}$ be the cumulative density function
% (or cumulative mass function if $\AgentDist(\cdot)$ is a discrete distribution)
% $F_{i,i'}(z) = \Prob[\AgV \sim \AgentDist]{\AgV \in \FirstTwo{i}{i'}
%   \mbox{ and } (\ArmV{i}-\ArmV{i'}) \cdot \AgV \leq z}$
% of $(\ArmV{i}-\ArmV{i'}) \cdot \AgV$,
% on $\AgV \in \Omega_{i,i^{'}}$.
% In words, $F_{i,i'}$ is the distribution of the \emph{strength} of the
% preference of a random agent for arm $i$ over arm $i'$

% $\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
% and
% $ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$

\dkedit{Let \AlmostTied{z} be the cumulative density function
(or cumulative mass function if $\AgentDist(\cdot)$ is a discrete distribution)
$\AlmostTied{z} = \Prob[\AgV \sim \AgentDist]{%
  (\ArmV{\Best{\AgV}}-\ArmV{\Second{\AgV}}) \cdot \AgV \leq z}$.
In words, \AlmostTied{z} is the CDF of the \emph{strength} of the
preference of a random agent for his best arm over his second-best arm.}

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exists a constant \TieDensity such that for all $z \in \R^+$,
%for all pairs $i,i'$ of arms and 
$\AlmostTied{z} \leq \TieDensity \cdot z$.
\end{assumption}

\dkedit{\paragraph{Role of parameters:}
Our problem setting is characterized by a fairly large number of
parameters.
Of these, we consider \ARMNUM, $\MinProb \leq 1/\ARMNUM$ and $T$ to be
the key parameters, while $\sigma, \Diam, d, \TieDensity$ should be
considered constant.
We will keep track of these constants throughout most of our proofs,
but report final results in terms of only the three key parameters
(except where we illustrate a specific point).}

\dkcomment{Stuff to discussion in conclusions:
\begin{enumerate}
\item Arms that are noone's best.
\item Other utilities besides inner products.
\item Other noise distributions.
\item Better bounds.
\end{enumerate}
}
